---
title: "Project 1"
author: "Vivi Feathers, Li-Chia Chang"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: show
    theme: readable
    df_print: tibble
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE,  warning = FALSE)
```

# Part 1 - Data Processing

## First Step
EDU01a data processing.

Read in the EDU01a data as sheet1 using the code below.

```{r edi01a_readin}
library(readr)
sheet1 <- read_csv("https://www4.stat.ncsu.edu/~online/datasets/EDU01a.csv")
sheet1
```

### 1. Select columns 
Select Area_name column and rename as area_name; select STCOU column and any column that ends in "D" and save it into `sheet1_select`.

```{r edi01a_select}
library(tidyverse)
sheet1_select <- sheet1 %>%
                   rename("area_name" = Area_name) %>% ## rename Area_name to area_name
                   select(area_name, STCOU, ends_with("D")) ## select columns as required
sheet1_select
```

### 2. Convert `sheet1_select` into long format 
Convert `sheet1_select` into long format where each row has one enrollment value for `area_name`, save it into `sheet1_convert`.

```{r edi01a_convert}
sheet1_convert <- sheet1_select %>%
## convert to long format and create 2 new columns
                    pivot_longer(cols = 3:12, names_to = "survey_id", values_to = "enrollment")
sheet1_convert
```

### 3. Create 2 new columns
Split the survey_id column into 2 new columns, one has the last 2 digits to represent year in numeric format, one has the first 3 characters and 4 digits to represent measurement. Save the output into `sheet1_newcol`.

```{r edi01a_newcol}
sheet1_newcol <- sheet1_convert %>%
## get the 8th and 9th sub string from survey_id, concatenate with "19" for year format, convert character to numeric, then get the first 7 sub string from survey_id, put them into new columns 
                   mutate(year = as.numeric(paste0(if_else(as.numeric(substr(sheet1_convert$survey_id, 8, 9)) > 10, "19", "20"), substr(sheet1_convert$survey_id, 8, 9))), measuement = substr(sheet1_convert$survey_id, 1, 7)) 
sheet1_newcol
```

### 4. Create two data sets.
Subset `sheet1_newcol` into 2 data sets: one contains only non-county data, add a class to the tibble called state, save it in `sheet1_noncounty`; and one contains only county level data, add a class to the tibble called county, save it in `sheet1_county`.

```{r edi01a_subset}
sheet1_county <- sheet1_newcol %>%
## find records that have area_name following the "County Name, DD" pattern
                   subset(grepl(pattern = ", \\w\\w", sheet1_newcol$area_name))
## add class called "county"
                   class(sheet1_county) <- c("county", class(sheet1_county))
sheet1_county

## find records that do not have area_name value in sheet1_county area_name values 
sheet1_noncounty <- sheet1_newcol[! sheet1_newcol$area_name %in% sheet1_county$area_name, ]
## add class called "state"
                   class(sheet1_noncounty) <- c("state", class(sheet1_noncounty))
sheet1_noncounty
```

### 5. Create state variable for the county level tibble
Use subset function to obtain the state name and output into a new column 
```{r edi01a_state}
sheet1_county <- sheet1_county %>%
## store the state info into a new column by using sub function and get the sub string after "," from column area_name
                   mutate(state = sub(".+, (.+)", "\\1", sheet1_county$area_name))
sheet1_county
```

### 6. Create a division variable for the non-county level tibble
According to [WIKIPEDIA](https://en.wikipedia.org/wiki/List_of_regions_of_the_United_States), find the division name for each state using `if_else` function.
```{r edi01a_division}
## check the distinct value in area_name
#unique(sheet1_noncounty$area_name)

sheet1_noncounty <- sheet1_noncounty %>%
                      mutate(division = if_else(area_name == "UNITED STATES", "ERROR",
                                                if_else(area_name %in% c("CONNECTICUT", "MAINE", "MASSACHUSETTS", "NEW HAMPSHIRE", "RHODE ISLAND", "VERMONT"), "NEW ENGLAND",
                                                        if_else(area_name %in% c("ILLINOIS", "INDIANA", "MICHIGAN", "OHIO", "WISCONSIN"), "EAST NORTH CENTRAL",
                                                                if_else(area_name %in% c("IOWA", "KANSAS", "MINNESOTA", "MISSOURI", "NEBRASKA", "NORTH DAKOTA", "SOUTH DAKOTA"), "WEST NORTH CENTRAL",
                                                                        if_else(area_name %in% c("NEW JERSEY", "NEW YORK", "PENNSYLVANIA"), "MIDDLE ATLANTIC", 
                                                                                if_else(area_name %in% c("DELAWARE", "FLORIDA", "GEORGIA", "MARYLAND", "NORTH CAROLINA", "SOUTH CAROLINA", "VIRGINIA", "DISTRICT OF COLUMBIA", "District of Columbia", "WEST VIRGINIA"), "SOUTH ATLANTIC", 
                                                                                        if_else(area_name %in% c("ALABAMA", "KENTUCKY", "MISSISSIPPI", "TENNESSEE"), "EAST SOUTH CENTRAL", 
                                                                                                if_else(area_name %in% c("ARKANSAS", "LOUISIANA", "OKLAHOMA", "TEXAS"), "WEST SOUTH CENTRAL",
                                                                                                        if_else(area_name %in% c("ARIZONA", "COLORADO", "IDAHO", "MONTANA", "NEVADA", "NEW MEXICO", "UTAH", "WYOMING"), "MOUNTAIN",
                                                                                                                if_else(area_name %in% c("ALASKA", "CALIFORNIA", "HAWAII", "OREGON", "WASHINGTON"), "PACIFIC", "")))))))))))
sheet1_noncounty
```

## Requirements
Now we want to repeat the above process for the 2nd component of the data set. Instead of than copying and pasting from previous code, we want to write functions that do above steps then create a wrapped function to do everything in one function call.

### 1. Write one function that does steps 1 & 2 above
Combine step 1 and 2 into one function also give an optional argument with a default value that allows the user to specify the name of the column representing the value (enrollment for these data sets).
```{r steponetwo}
##`df1` is for data frame name and `col_name` is for defining the column name which stores the enrollment values`
step_one_two <- function(df1, var_name) {
                                   df1 %>%
                                     rename("area_name" = Area_name) %>% ## rename Area_name to area_name
                                     select(area_name, STCOU, ends_with("D")) %>% ## select columns as required
                                     pivot_longer(cols = 3:12, names_to = "survey_id", values_to = var_name) ## convert to long format and create 2 new columns
}
#step_one_two (sheet1, "enroll")
```

### 2. Write one function that does steps 3 above
Write one function that takes in the output from step 1 & 2, and creates a new tibble in step 3
```{r stepthree}
##`df2` is for data frame name
step_three <- function(df2) {
                                  df2 %>%
## get the 8th and 9th sub string from survey_id, concatenate with "19" for year format, convert character to numeric, then get the first 7 sub string from survey_id, put them into new columns 
                                    mutate(year = as.numeric(paste0(if_else(as.numeric(substr(df2$survey_id, 8, 9)) > 10, "19", "20"), substr(df2$survey_id, 8, 9))), measuement = substr(df2$survey_id, 1, 7)) 
}
#step_three (sheet1_convert)
```

### 3. Write one function that does steps 5 above
Write one function that takes in the output from step 4 and creates a new tibble in step 5.
```{r stepfive}
##`df4` is for data frame name
step_five <- function(df4) {
              sheet1_county_final <<- df4 %>%
## store the state info into a new column by using sub function and get the sub string after "," from column area_name
                                   mutate(state = sub(".+, (.+)", "\\1", df4$area_name))
              
}
#step_five (sheet1_county)
```

### 4. Write one function that does steps 6 above
Write one function that takes in the output from step 4 and creates a new tibble in step 6.
```{r stepsix}
##`df5` is for data frame name
step_six <- function(df5) {
              sheet1_noncounty_final <<- df5 %>%
                                     mutate(division = if_else(area_name == "UNITED STATES", "ERROR",
                                                        if_else(area_name %in% c("CONNECTICUT", "MAINE", "MASSACHUSETTS", "NEW HAMPSHIRE", "RHODE ISLAND", "VERMONT"), "NEW ENGLAND",
                                                         if_else(area_name %in% c("ILLINOIS", "INDIANA", "MICHIGAN", "OHIO", "WISCONSIN"), "EAST NORTH CENTRAL",
                                                          if_else(area_name %in% c("IOWA", "KANSAS", "MINNESOTA", "MISSOURI", "NEBRASKA", "NORTH DAKOTA", "SOUTH DAKOTA"), "WEST NORTH CENTRAL",
                                                           if_else(area_name %in% c("NEW JERSEY", "NEW YORK", "PENNSYLVANIA"), "MIDDLE ATLANTIC", 
                                                            if_else(area_name %in% c("DELAWARE", "FLORIDA", "GEORGIA", "MARYLAND", "NORTH CAROLINA", "SOUTH CAROLINA", "VIRGINIA", "DISTRICT OF COLUMBIA", "District of Columbia", "WEST VIRGINIA"), "SOUTH ATLANTIC", 
                                                             if_else(area_name %in% c("ALABAMA", "KENTUCKY", "MISSISSIPPI", "TENNESSEE"), "EAST SOUTH CENTRAL", 
                                                              if_else(area_name %in% c("ARKANSAS", "LOUISIANA", "OKLAHOMA", "TEXAS"), "WEST SOUTH CENTRAL",
                                                               if_else(area_name %in% c("ARIZONA", "COLORADO", "IDAHO", "MONTANA", "NEVADA", "NEW MEXICO", "UTAH", "WYOMING"), "MOUNTAIN",
                                                                if_else(area_name %in% c("ALASKA", "CALIFORNIA", "HAWAII", "OREGON", "WASHINGTON"), "PACIFIC", "")))))))))))
              #return (sheet1_noncounty_final)
}
#step_six (sheet1_noncounty)
```

### 5. Write one function that does steps 4 above
Write one function that takes in the output from step 3 and creates the two tibbles in step 4, calls step_five and step_six functions then returns two final tibbles.
```{r stepfour}
##`df3` is for data frame name, save the output into `sheet1_county1` and "sheet1_noncounty1"
step_four <- function(df3) {
               sheet1_county1 <- df3 %>%
## find records that have area_name following the "County Name, DD" pattern
                                   subset(grepl(pattern = ", \\w\\w", df3$area_name))
## add class called "county"
                                   class(sheet1_county1) <- c("county", class(sheet1_county1))
## call step_five function
                                   step_five(sheet1_county1)

## find records that do not have area_name value in sheet1_county area_name values 
               sheet1_noncounty1 <- df3[! df3$area_name %in% sheet1_county1$area_name, ]
## add class called "state"
                                      class(sheet1_noncounty1) <- c("state", class(sheet1_noncounty1))
## call step_six function
                                      step_six(sheet1_noncounty1)
            return (list(sheet1_county_final, sheet1_noncounty_final))
}
#step_four (sheet1_newcol)
```
### 6. Write one function that reads in the .csv file
Create a function that takes in the URL of a .csv file
```{r stepzero}
##`url` is for the data source url
step_zero <- function(url) {
               read_csv(url)
}
```

### 7. Create the final wrapped function
Create a wrapper function calls all the functions written above, then returns the two tibbles.
```{r wrap}
##`url` is for the data source url, `var_name` is an optional argument for the variable name
wrap_up <- function(url, var_name = "enroll") {
           step0 <- step_zero(url)
           step_1_2 <- step_one_two (step0, var_name)
           step3 <- step_three (step_1_2)
           step_four (step3)
}
## RUN IT!!!
wrap_up ("https://www4.stat.ncsu.edu/~online/datasets/EDU01b.csv")
```